---
title: "p8105_hw3_ys3875"
author: "Yanhao Shen"
date: "2024-10-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(janitor)
```

Q2.
```{r}
demographic_data =
  read_csv("nhanes_covar.csv", na = c('NA', ' ', '.'), col_names = TRUE, skip = 4) |>
  drop_na() |>
  janitor::clean_names() |>
  filter(age>=21) |>
  mutate(
    sex=
      case_match(
        sex,
        1~ 'male',
        2~ 'female'),
    education=
      case_match(
        education,
        1~ 'Less than high school',
        2~ 'High school equivalent',
        3~ 'More than high school'),
    sex = as.factor(sex),
    education = as.factor(education)
  )

accelerometer_data = 
  read_csv("nhanes_accel.csv", na=c('NA',' ','.')) |>
  drop_na() |>
  janitor::clean_names()
```
```{r}
dd_table <- demographic_data |>
  group_by(education,sex) |>
  summarise(count=n(), .groups = 'drop')|>
  spread(key = sex, value=count,fill = 0)
dd_table

dd_plot <- demographic_data |>
  ggplot(aes(x=education, y=age, fill=sex))+
  geom_violin()+
  theme_minimal()
dd_plot
```

Relatively, there are more men going to high school than women, for less than and more that high school, the number are pretty close. \
For both man and women, the age lies averagely for high school equivalent. There are less young energy(<30) for less than high school, while for more than high school, the condition is exactly the opposite.
```{r}
total_activity_data <- 
  accelerometer_data |>
  rowwise() |>
  mutate(total_activity=sum(c_across(min1:min1440),na.rm=TRUE)) |>
  select(seqn, total_activity) |>
  left_join(demographic_data,by='seqn') |>
  na.omit()

ggplot(total_activity_data, aes(x=age, y=total_activity, color=sex))+
  geom_point()+
  geom_smooth(method = 'loess', se=FALSE)+
  facet_wrap(.~education)+
  theme_minimal()
```

We can see from the plots that no matter what education level for both man and woman, their total activital tend to get smaller as the age goes up.
```{r}
acc_data <- accelerometer_data |>
  pivot_longer(cols = starts_with('min'),
               names_to = 'minutes',
               values_to = 'activity') |>
  left_join(demographic_data,by='seqn') |>
  na.omit()
ggplot(acc_data, aes(x=minutes,y=activity,color=sex))+
         geom_line()+
         geom_smooth(se=FALSE)+
         facet_wrap(.~education)+
         theme_minimal()
```

People are not willing to do activities during noon, and those who are more than high school tend to do more compared to the other two groups.

Q3.
```{r}
Jan_2020_Citi <- read_csv("citibike/Jan 2020 Citi.csv", na = c('NA', ' ', '.')) |>
  drop_na() |>
  janitor::clean_names() |>
  mutate(year=2020, month='January')

July_2020_Citi <- read_csv("citibike/July 2020 Citi.csv",na = c('NA', ' ', '.')) |>
  drop_na() |>
  janitor::clean_names() |>
  mutate(year=2020, month='July')

Jan_2024_Citi <- read_csv("citibike/Jan 2024 Citi.csv",na = c('NA', ' ', '.')) |>
  drop_na() |>
  janitor::clean_names() |>
  mutate(year=2024, month='January')

July_2024_Citi <- read_csv("citibike/July 2024 Citi.csv",na = c('NA', ' ', '.')) |>
  drop_na() |>
  janitor::clean_names() |>
  mutate(year=2024, month='July')

Q3_data_all <- bind_rows(Jan_2020_Citi,July_2020_Citi,Jan_2024_Citi,July_2024_Citi) |>
  janitor::clean_names()

  
```
The datasets show the data of bike usage.
```{r}
total_rides_data <- Q3_data_all |>
  group_by(year,month,member_casual) |>
  summarise(total_rides=n(), .groups = 'drop')
```
People are more willing to ride during July.
```{r}
top_5_starting_station_July2024 <- July_2024_Citi |>
  group_by(start_station_name) |>
  summarise(total_rides=n(),.groups = 'drop') |>
  arrange(desc(total_rides)) |>
  slice(1:5)
```

```{r}
Q3_data_all_median <- Q3_data_all |>
  group_by(year, month, weekdays) |>
  summarise(median_weekdays=median(duration),.groups = 'drop')

ggplot(Q3_data_all_median, 
       aes(x=weekdays, y=median_weekdays,
           color=year)) +
  geom_point()+
  geom_line()+
  theme_minimal()
```

The plot tells us that people are more willing to go riding during weekends.
```{r}
data_2024 <- Q3_data_all |>
  filter(year==2024)

ggplot(data_2024,aes(x=duration, y=rideable_type))+
  geom_point()+
  facet_grid(member_casual~month) +
  theme_minimal()
```

For casual people, they use electric bike more often compared to the members.
